The interactions of rational, pragmatic agents\\ lead to efficient language structure and use

### 2019-02-01 paper/paper.Rmd
BP final read edits

We adopt a different approach in this work in two ways. First we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems. Second we demonstrate how assumptions about language-users impacts whether language properties are used efficiently. In this way, we bridge questions of design and use in a single framework.\par

@@

We provide a case study for this approach, in which functionalist regularities emerge from the dynamics of pragmatic communication. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanadum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction. Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency. Further, we show how ambiguity is \textit{used} efficiently during local, in-the-moment interactions. The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par

@@

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the number of context-sizes (1-4). Optimal language under CE objective (red). Speaker-optimal (blue). Listener-optimal (green). Error bars represent 95 percent confidence intervals. Panel (B), example CE-optimal language (ambiguous) under in a four-condition context. Panel (C), example CE-optimal language (unambiguous) in a single-condition context."}
img <- png::readPNG("figs/fig1.png")
grid::grid.raster(img)
```

@@

In Simulation 1 we showed that efficiency defined in terms of pragmatic agents leads to a preference for languages that contain ambiguity. In Simulation 2 we assume a single fixed language $\ell$, which contains ambiguity, and instead vary the types of agents using $\ell$. We will show that efficient *use* of ambiguity depends on an agent's ability to use context for disambiguation. This instantiates an on-line perspective on the communicative function of ambiguity theory. More generally, it demonstrates how the reference game framework can be used to asses questions of efficient language use (current simulation) as well as language design (Simulation 1).\par

@@

Figure 2 (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse. The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. Panel (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs (see panel B). While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par

@@

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{link}.}}} \vspace{1em}
\noindent

>>>>

To preview, we diverge from this approach in two ways. First we derive a measure of efficiency from first principles of \textit{efficient language use} and show how this can lead to communicatively efficient systems. Second we demonstrate how assumptions about language-users impacts whether language properties are used efficiently. In this way, we bridge questions of design and use in a single framework.\par

@@

We provide a case study for this approach, in which functionalist regularities emerge from the dynamics of pragmatic communication. We choose a property of languages that could, in principle, vary freely, but shows strong regularities across languages. The explanadum is why this regularity holds. We examine ambiguity as our property, extending ideas by Piantadosi et al. [-@Piantadosi2011a]. We define a novel measure of efficiency that depends on the interactional behavior of speaker and listener agents. We adopt the reference game as our primary unit of interaction and model language users with the Rational Speech Act (RSA) framework -- a computional model of language use, which is nicely attested by experimental data on interaction. Using these basic ingredients, we show that the property of interest (ambiguity) is prevalent in languages that optimize our measure of efficiency. Further, we show \textit{how} ambiguity is used efficiently during local, in-the-moment interactions. The contributions of this work are twofold -- we derive a novel objective for measuring linguistic efficiency and also show how the reference game framework, in combination with formal models of communication, can be used to connect ideas about system-level efficiencies to in-the-moment language use.\par

@@

```{r plot-optimal-langs, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2.8, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the context-size (1-4) in each condition. Optimal language under CE objective (red). Speaker-optimal (blue). Listener-optimal (green). Error bars represent 95 percent confidence intervals. Panel (B), example CE-optimal language (ambiguous) under in a four-condition context. Panel (C), example CE-optimal language (unambiguous) in a single-condition context."}
img <- png::readPNG("figs/fig1.png")
grid::grid.raster(img)
```

@@

In Simulation 1 we showed that efficiency defined in terms of pragmatic agents leads to a preference for languages that contain ambiguity. In Simulation 2 we assume a single fixed language $\ell$, which contains ambiguity, and instead vary the types of agents using $\ell$. We will show that efficient *use* of ambiguity depends on an agent's ability to use context for disambiguation. This instantiates an on-line perspective on the communicative function of ambiguity theory. More generally, it demonstrates how we can assess both questions of efficient use (current simulation) as well as design (previous simulation) in the same framework.\par

@@

Figure 3, panel (A) shows the empirical probability that a speaker uses an ambiguous utterance as a function of discourse position. The \textit{No pragmatic} baseline uses ambiguous utterances frequently and at a constant rate over the discourse. The \textit{Partial pragmatic} baseline avoids ambiguous utterances entirely. But, the \textit{Full pragmatic} model avoids ambiguous material only at the start of the discourse, employing it increasingly as the discourse proceeds. Panel (C) tracks our CE objective for each model over the discourse. Note that the objective decreases for all three models, primarily driven by the listener updating his belief about the actual topic ($P(C=c_{current}|D)$). However, the objective declines more quickly under the \textit{Full} and \textit{Partial} pragmatic speakers as listener agents are better able to infer the correct context. Additionally, the difference in CE between the \textit{Full}- and \textit{Partial pragmatic} models at the end of the discourse is driven by the reduction in speaker costs (see panel B). While speaker effort remains constant in both \textit{No pragmatic} and \textit{Partial pragmatic} baselines, effort declines in the \textit{Full pragmatic} model as she increasingly relies on ambiguous material later in the discourse.\par

@@

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{repository}.}}} \vspace{1em}
\noindent

### 2019-02-01 paper/paper.tex
BP final read edits

We provide a case study for this approach, in which functionalist
regularities emerge from the dynamics of pragmatic communication. We
choose a property of languages that could, in principle, vary freely,
but shows strong regularities across languages. The explanadum is why
this regularity holds. We examine ambiguity as our property, extending
ideas by Piantadosi et al. (2011). We define a novel measure of
efficiency that depends on the interactional behavior of speaker and
listener agents. We adopt the reference game as our primary unit of
interaction and model language users with the Rational Speech Act (RSA)
framework -- a computional model of language use, which is nicely
attested by experimental data on interaction. Using these basic
ingredients, we show that the property of interest (ambiguity) is
prevalent in languages that optimize our measure of efficiency. Further,
we show how ambiguity is \textit{used} efficiently during local,
in-the-moment interactions. The contributions of this work are twofold
-- we derive a novel objective for measuring linguistic efficiency and
also show how the reference game framework, in combination with formal
models of communication, can be used to connect ideas about system-level
efficiencies to in-the-moment language use.\par

@@

\caption[Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity]{Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the number of context-sizes (1-4). Optimal language under CE objective (red). Speaker-optimal (blue). Listener-optimal (green). Error bars represent 95 percent confidence intervals. Panel (B), example CE-optimal language (ambiguous) under in a four-condition context. Panel (C), example CE-optimal language (unambiguous) in a single-condition context.}\label{fig:plot-optimal-langs}
\end{figure*}
\end{CodeChunk}

@@

In Simulation 1 we showed that efficiency defined in terms of pragmatic
agents leads to a preference for languages that contain ambiguity. In
Simulation 2 we assume a single fixed language \(\ell\), which contains
ambiguity, and instead vary the types of agents using \(\ell\). We will
show that efficient \emph{use} of ambiguity depends on an agent's
ability to use context for disambiguation. This instantiates an on-line
perspective on the communicative function of ambiguity theory. More
generally, it demonstrates how the reference game framework can be used
to asses questions of efficient language use (current simulation) as
well as language design (Simulation 1).\par

@@

Figure 2 (A) shows the empirical probability that a speaker uses an
ambiguous utterance as a function of discourse position. The
\textit{No pragmatic} baseline uses ambiguous utterances frequently and
at a constant rate over the discourse. The \textit{Partial pragmatic}
baseline avoids ambiguous utterances entirely. But, the
\textit{Full pragmatic} model avoids ambiguous material only at the
start of the discourse, employing it increasingly as the discourse
proceeds. Panel (C) tracks our CE objective for each model over the
discourse. Note that the objective decreases for all three models,
primarily driven by the listener updating his belief about the actual
topic (\(P(C=c_{current}|D)\)). However, the objective declines more
quickly under the \textit{Full} and \textit{Partial} pragmatic speakers
as listener agents are better able to infer the correct context.
Additionally, the difference in CE between the \textit{Full}- and
\textit{Partial pragmatic} models at the end of the discourse is driven
by the reduction in speaker costs (see panel B). While speaker effort
remains constant in both \textit{No pragmatic} and
\textit{Partial pragmatic} baselines, effort declines in the
\textit{Full pragmatic} model as she increasingly relies on ambiguous
material later in the discourse.\par

@@

\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{link}.}}}
\vspace{1em} \noindent

>>>>

We provide a case study for this approach, in which functionalist
regularities emerge from the dynamics of pragmatic communication. We
choose a property of languages that could, in principle, vary freely,
but shows strong regularities across languages. The explanadum is why
this regularity holds. We examine ambiguity as our property, extending
ideas by Piantadosi et al. (2011). We define a novel measure of
efficiency that depends on the interactional behavior of speaker and
listener agents. We adopt the reference game as our primary unit of
interaction and model language users with the Rational Speech Act (RSA)
framework -- a computional model of language use, which is nicely
attested by experimental data on interaction. Using these basic
ingredients, we show that the property of interest (ambiguity) is
prevalent in languages that optimize our measure of efficiency. Further,
we show \textit{how} ambiguity is used efficiently during local,
in-the-moment interactions. The contributions of this work are twofold
-- we derive a novel objective for measuring linguistic efficiency and
also show how the reference game framework, in combination with formal
models of communication, can be used to connect ideas about system-level
efficiencies to in-the-moment language use.\par

@@

\caption[Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity]{Panel (A) Vertical axis shows the proportion of optimal languages containing ambiguity. Horizontal axis shows the context-size (1-4) in each condition. Optimal language under CE objective (red). Speaker-optimal (blue). Listener-optimal (green). Error bars represent 95 percent confidence intervals. Panel (B), example CE-optimal language (ambiguous) under in a four-condition context. Panel (C), example CE-optimal language (unambiguous) in a single-condition context.}\label{fig:plot-optimal-langs}
\end{figure*}
\end{CodeChunk}

@@

In Simulation 1 we showed that efficiency defined in terms of pragmatic
agents leads to a preference for languages that contain ambiguity. In
Simulation 2 we assume a single fixed language \(\ell\), which contains
ambiguity, and instead vary the types of agents using \(\ell\). We will
show that efficient \emph{use} of ambiguity depends on an agent's
ability to use context for disambiguation. This instantiates an on-line
perspective on the communicative function of ambiguity theory. More
generally, it demonstrates how we can assess both questions of efficient
use (current simulation) as well as design (previous simulation) in the
same framework.\par

@@

Figure 3, panel (A) shows the empirical probability that a speaker uses
an ambiguous utterance as a function of discourse position. The
\textit{No pragmatic} baseline uses ambiguous utterances frequently and
at a constant rate over the discourse. The \textit{Partial pragmatic}
baseline avoids ambiguous utterances entirely. But, the
\textit{Full pragmatic} model avoids ambiguous material only at the
start of the discourse, employing it increasingly as the discourse
proceeds. Panel (C) tracks our CE objective for each model over the
discourse. Note that the objective decreases for all three models,
primarily driven by the listener updating his belief about the actual
topic (\(P(C=c_{current}|D)\)). However, the objective declines more
quickly under the \textit{Full} and \textit{Partial} pragmatic speakers
as listener agents are better able to infer the correct context.
Additionally, the difference in CE between the \textit{Full}- and
\textit{Partial pragmatic} models at the end of the discourse is driven
by the reduction in speaker costs (see panel B). While speaker effort
remains constant in both \textit{No pragmatic} and
\textit{Partial pragmatic} baselines, effort declines in the
\textit{Full pragmatic} model as she increasingly relies on ambiguous
material later in the discourse.\par

@@

\fbox{\parbox[b][][c]{7.3cm}{\centering {Code and SI are available at this \href{https://github.com/benpeloquin7/zipf_principles/}{repository}.}}}
\vspace{1em} \noindent
